---
title: "Lobster Report"
Authors: "Soraida Garcia and Beatriz Mejia Mercado"
format: html
---

## Exercise: Explore, Clean, and Wrangle Data


### Abstract

Data on abundance, size and fishing pressure of California spiny lobster (Panulirus interrupt) are collected along the mainland coast of the Santa Barbara Channel. Spiny lobsters are an important predator in giant kelp forests off southern California. Two SBC LTER study reefs are located in or near the California Fish and Game Network of Marine Protected Areas (MPA), Naples and Isla Vista, both established as MPAs on 2012-01-01. MPAs provide a unique opportunity to investigate the effects of fishing on kelp forest community dynamics. Sampling began in 2012 and is ongoing.

### Set up

```{r}
#| warning: false
#| message: false

library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
```

### Read the data - Abundance

```{r}
lobster_abundance <- read.csv("../Beatriz-Soraida/data/Lobster_Abundance_All_Years_20220829.csv")
```

### Exploring the data
```{r}
lobster_abundance <- lobster_abundance %>% 
    mutate(SIZE_MM = na_if(SIZE_MM, -99999))

not_napl <- lobster_abundance %>% 
    filter(SITE != "NAPL")

aque_70mm <- lobster_abundance %>% 
    filter(SITE == "AQUE", SIZE_MM >= 70)

```


## Exercise: Data Visualization

```{r}
#| warning: false
#| message: false

ggplot(data = lobster_abundance, aes(x = SIZE_MM)) +
    geom_histogram() +
    facet_wrap(~SITE)


lobsters_summarize <- lobster_abundance %>% 
  group_by(SITE, YEAR) %>% 
  summarize(COUNT = n()) %>%
  mutate(YEAR=as.factor(YEAR))  ##this eliminates the decimals
  

lobster_size_lrg <- lobster_abundance %>% 
    filter(YEAR %in% c(2019, 2020, 2021)) %>% 
    mutate(SIZE_BIN = if_else(SIZE_MM <= 70, true = "small", false = "large")) %>% 
    group_by(SITE, SIZE_BIN) %>%
    summarize(COUNT = n()) %>%
    drop_na()
```


### Bar plot
```{r}
ggplot(data = lobster_size_lrg, aes(x = SITE, y = COUNT, fill = SIZE_BIN)) +
    geom_col()
```

### Dodged bar plot
```{r}
ggplot(data = lobster_size_lrg, aes(x = SITE, y = COUNT, fill = SIZE_BIN)) +
    geom_col(position = "dodge")


```


# Line and point plot

```{r}
#| warning: false
#| message: false

ggplot(data = lobsters_summarize, aes(x = YEAR, y = COUNT)) +
  geom_point(aes(color = SITE)) +
  geom_line(aes(color = SITE)) + 
  facet_wrap( ~ SITE,
                scales = "free_y") + 
  theme_light() +
    theme(legend.position = "bottom",
          legend.title = element_blank()) +
  theme(axis.text.x = element_text(angle = 45,
                                     vjust = 0.5),
        axis.title.x = element_text(vjust = -2)) +
   ggtitle("Lobster Abundance")
```



### Read the data - Trap counts

```{r}
#| warning: false

lobster_traps <- read.csv("../Beatriz-Soraida/data/Lobster_Trap_Counts_All_Years_20210519.csv")
```

### Exploring the data

```{r}
glimpse(lobster_traps)


summary(lobster_traps)

```

### Convert missing values using `mutate()` and `na_if()`

```{r}
lobster_traps <- lobster_traps %>% 
  mutate(TRAPS = na_if(TRAPS, -99999))
```


### `filter()` practice

```{r}
not_napl <- lobster_traps %>% 
  filter(SITE != "NAPL")
```


```{r}
Carp_20_traps <- lobster_traps %>% 
  filter(SITE == "CARP", TRAPS > 20)
```


## Exercise: Data Visualization

```{r}

ggplot(data = lobster_traps, aes(x = TRAPS))+
  geom_histogram(bins = 15, na.rm = TRUE)+
  facet_wrap( ~ YEAR)

```

```{r}
#| message: false

lobsters_traps_summarize <- lobster_traps %>% 
  group_by(SITE, YEAR) %>% 
  summarize(TOTAL_TRAPS = sum(TRAPS, na.rm = TRUE))
```
### line plot

```{r}
ggplot(data = lobsters_traps_summarize, aes(x = YEAR,
                                          y = TOTAL_TRAPS))+
  geom_line(aes(color = SITE))
```

### line and point plot

```{r}
ggplot(data = lobsters_traps_summarize, aes(x = YEAR,
                                            y = TOTAL_TRAPS))+
  geom_point(aes(color = SITE))+
  geom_line(aes(color = SITE))
```
### Bar graph

```{r}
#| message: false

lobster_traps_fishing_pressure <- lobster_traps %>% 
  filter(YEAR %in% c(2019, 2020, 2021)) %>% 
  mutate(FISHING_PRESSURE = if_else(TRAPS >=8,
                                    true = "high",
                                    false = "low")) %>% 
  group_by(SITE, FISHING_PRESSURE) %>%
  summarize(COUNT = n()) %>% 
  drop_na()
             
```
### Bar plot

```{r}
ggplot(data = lobster_traps_fishing_pressure,
       aes(x = SITE,
           y = COUNT,
           fill = FISHING_PRESSURE))+
  geom_col()+
  scale_fill_viridis_d()+
  theme(legend.position = "bottom")
```

### Dodged bar plot

```{r}
ggplot(data = lobster_traps_fishing_pressure,
       aes(x = SITE,
           y = COUNT,
           fill = FISHING_PRESSURE))+
  geom_col(position = "dodge")+
  scale_fill_viridis_d()+
  theme(legend.position = "bottom")
```




## Citation 

Santa Barbara Coastal LTER, D. Reed, and R. Miller. 2022. SBC LTER: Reef: Abundance, size and fishing effort for California Spiny Lobster (Panulirus interruptus), ongoing since 2012 ver 8. Environmental Data Initiative. https://doi.org/10.6073/pasta/25aa371650a671bafad64dd25a39ee18 (Accessed 2023-10-04).




