---
title: "Collaborator Analysis"
Author: Beatriz Mejia-Mercado
format: html
---

## Exercise: Explore, Clean, and Wrangle Data

### About this data

Data on abundance, size and fishing pressure of California spiny lobster (Panulirus interruptus) along the mainland coast of the Santa Barbara Channel from 2012-2021.

### Setup
```{r}
#| warning: false
#| comment: false

library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
```

### Read the data

```{r}
lobster_traps <- read.csv("../data/Lobster_Trap_Counts_All_Years_20210519.csv")
```

### Exploring the data

```{r}
glimpse(lobster_traps)


summary(lobster_traps)

```

### Convert missing values using `mutate()` and `na_if()`

```{r}
lobster_traps <- lobster_traps %>% 
  mutate(TRAPS = na_if(TRAPS, -99999))
```


### `filter()` practice

```{r}
not_napl <- lobster_traps %>% 
  filter(SITE != "NAPL")
```


```{r}
Carp_20_traps <- lobster_traps %>% 
  filter(SITE == "CARP", TRAPS > 20)
```


## Exercise: Data Visualization

```{r}
ggplot(data = lobster_traps, aes(x = TRAPS))+
  geom_histogram()+
  facet_wrap( ~ YEAR)

```

```{r}
lobsters_traps_summarize <- lobster_traps %>% 
  group_by(SITE, YEAR) %>% 
  summarize(TOTAL_TRAPS = sum(TRAPS, na.rm = TRUE))
```
### line plot

```{r}
ggplot(data = lobsters_traps_summarize, aes(x = YEAR,
                                          y = TOTAL_TRAPS))+
  geom_line(aes(color = SITE))
```

### line and point plot

```{r}
ggplot(data = lobsters_traps_summarize, aes(x = YEAR,
                                            y = TOTAL_TRAPS))+
  geom_point(aes(color = SITE))+
  geom_line(aes(color = SITE))
```
### Bar graph

```{r}
lobster_traps_fishing_pressure <- lobster_traps %>% 
  filter(YEAR %in% c(2019, 2020, 2021)) %>% 
  mutate(FISHING_PRESSURE = if_else(TRAPS >=8,
                                    true = "high",
                                    false = "low")) %>% 
  group_by(SITE, FISHING_PRESSURE) %>%
  summarize(COUNT = n()) %>% 
  drop_na()
             
```
### Bar plot

```{r}
ggplot(data = lobster_traps_fishing_pressure,
       aes(x = SITE,
           y = COUNT,
           fill = FISHING_PRESSURE))+
  geom_col()
```

### Dodged bar plot

```{r}
ggplot(data = lobster_traps_fishing_pressure,
       aes(x = SITE,
           y = COUNT,
           fill = FISHING_PRESSURE))+
  geom_col(position = "dodge")
```

